Formulating the posterior for \tardis

### 2015-10-28 posterior/model.tex
[model.tex] Minor reformatting

%%% refs
\def \refeq#1{(\ref{eq:#1})}
\def \refsec#1{sec.~\ref{#1}}
\def \refSec#1{Sec.~\ref{#1}}
\def \refapp#1{app.~\ref{#1}}
\def \reffig#1{fig.~\ref{fig:#1}}

@@

\section{Input and Output}

\subsubsection*{What we are given}
\begin{itemize}
\item data $\data \sim \expin =$ set of input values from
  telescope for each frequency bin $i=1 \dots n_b$;
\item \tardis{} depends on the parameters $\vecth$ and
  produces $n_\real$ real photon packets that leave the supernova and
  $n_\virt$ virtual photon packets. When we bin the packets in the same
  frequency bins as \expin, the sum of luminosities in bin $i$ is
  $\lumi^\real$ and $\lumi^\virt$.
\end{itemize}
I try to be mostly consistent with standard notation in statistics and
denote random variables---eventually they will be integrated out---by
capital letters and numbers that are observed by lower case.

@@

% The overall goal is to extract (or maximize) the posterior of
% $\vecth$, the parameters that describe the physics behind
% supernovae. Bayes' theorem is
% \begin{align}
%   \label{eq:post}
%   P(\vecth | \mathcal{D}) &\propto P(\mathcal{D} | \vecth) P_0(\vecth)
% \end{align}
\section{Likelihood}

@@

Given the KDE estimate, we could compute the distribution for the sum
of two variables as the convolution
\begin{align}
  \label{eq:sum-of-2}
  P(\Lumi | N_i=2, \vecth) = \int \rmdx{l} P(\Lumi -l | N_i=1, \vecth) P(l | N_i=1, \vecth).
\end{align}
For $N_i =k$, we would repeat the procedure convoluting $P(\cdot |
N_i=k-1, \vecth)$ and $P(\Lumi | N_i=1, \vecth)$ and each time we
would need a new KDE estimate and we would have to repeat the integral
for different values of $\Lumi$. This quickly gets out of hand and I
fear the errors due to KDE assumptions add up. In my attempts with
\texttt{mathematica}, the probability for many values of $\Lumi$ was
negative and the whole procedure is slow. Let's forget about it.

>>>>

%%% refs
\def \refeq#1{(\ref{eq:#1})}
\def \refsec#1{sec.~\ref{sec:#1}}
\def \refapp#1{app.~\ref{#1}}
\def \reffig#1{fig.~\ref{fig:#1}}

@@

\section{Overview}

\subsubsection*{Notation}
I try to be consistent with standard notation in statistics and
denote random variables by capital letters and numbers that are
observed by lower case.

\subsubsection*{What we are given}
\begin{itemize}
  \item $\expin =$ the set of input values from the telescope for each
  frequency bin $i=1 \dots n_b$. For the interpretation, see below in
  \refsec{likelihood}.
\item \tardis{} depends on the parameters $\vecth$ and
  produces $n_\real$ real photon packets that leave the supernova and
  $n_\virt$ virtual photon packets. When we bin the packets in the same
  frequency bins as \expin, the sum of luminosities in bin $i$ is
  $\lumi^\real$ and $\lumi^\virt$.
\end{itemize}

@@

% The overall goal is to extract (or maximize) the posterior of
% $\vecth$, the parameters that describe the physics behind
% supernovae. Bayes' theorem is
% \begin{align}
%   \label{eq:post}
%   P(\vecth | \mathcal{D}) &\propto P(\mathcal{D} | \vecth) P_0(\vecth)
% \end{align}
\section{Likelihood}\label{sec:likelihood}

@@

Given the KDE estimate, we could compute the distribution for the sum
of two variables as the convolution
\begin{align}
  \label{eq:sum-of-2}
  P(\Lumi | N_i=2, \vecth) = \int \rmdx{l} P(\Lumi -l | N_i=1, \vecth) P(l | N_i=1, \
vecth).
\end{align}
For $N_i =k$, we would repeat the procedure convoluting $P(\cdot |
N_i=k-1, \vecth)$ and $P(\Lumi | N_i=1, \vecth)$ and each time we
would need a new KDE estimate and we would have to repeat the integral
for different values of $\Lumi$. This quickly gets out of hand and I
fear the errors due to KDE assumptions add up. In my attempts with
\texttt{mathematica}, the probability for many values of $\Lumi$ was
negative and the whole procedure is slow. Let's forget about it.
